---
title: "Infectious_diseases"
author: "VC"
date: "5/13/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Infectious diseases are major concern in densely populated conditions. Understandig trends of infectious diseases prevented by vaccine can give a good insight on what and how the diseases spread and what conditions diseases depend on.
Here I have gathered infectious disease data of California from CDC, and demography data from 2010 census (https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?src=CF)
```{r }
##needed packages
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
###
cenus<- read.csv("DEC_10_SF1_GCTPH1.ST05_with_ann.csv",header=T)
head(cenus)
colnames(cenus)<- c("ID","ID2","Geog","Geoid1","Geoid2","Geographicarea",
                    "county","Population","Housing units","totalarea","waterarea",
                    "leandarea","popdens","housedens")

```

## Population size  per county

The following plot states the population size per county

```{r , echo=FALSE}
##
ggplot(cenus)+
  geom_col(aes(county,Population, fill=county))+
  ylim(0,1000000)+
  ggpubr::theme_pubr()+
  xlab("Counties")+
  ylab("Population")+
  theme(axis.text.x = element_text(size=16,hjust=45))+
  theme(axis.title.x = element_text(size=16))+
  theme(axis.text.y = element_text(size=16))+
  theme(axis.title.y = element_text(size=16))+
  labs(title = "Population of california counties")
```
## Disease distribution per county

The following plot states the population size per county
```{r , echo=FALSE, warning=FALSE}
##
dis<- read.csv("izb_odp_final_03262019.csv")
head(dis)
dis2<- dis %>% 
  spread(year,count)
head(dis2)    
ggplot(dis)+
  geom_line(aes(year,count, col=county))+
  facet_wrap(~disease,scales = "free")+
  ggpubr::theme_pubr()+
  xlab("Disease")+
  ylab("Number of cases")+
  theme(axis.text.x = element_text(size=16))+
  theme(axis.title.x = element_text(size=16))+
  theme(axis.text.y = element_text(size=16))+
  theme(axis.title.y = element_text(size=16))+
  labs(title = "Number of cases of infectious diseases in california counties")
```

#####Data wrangling to select data for 2010 only as census data is for 2010 only
```{r}
#split the data for county from census
head(cenus)
census<- cenus %>% 
  na.omit() %>% 
  separate("county",into=c("county","just "))
##merge by county
dem.dis<- full_join(census,dis,by="county")
head(dem.dis)
tail(dem.dis)
```

##data only for pertussis as it is the most common disease
```{r}
###does the cases depend on the population
dem.dis2<-dem.dis %>% filter(year==2010) %>% group_by(disease) %>% 
  select(county,disease,Population,count,popdens)
###pertusis
de.dis.2010.per<-dem.dis %>% 
  filter(year==2010) %>% 
  select(county,year,Population,disease,count,popdens) %>% 
 filter(disease== "Pertussis")
```
###GGplot with regression line to understand trenad with
#A. Population
```{r}
### relation between population and cases of pertusis
ggplot(de.dis.2010.per, aes(Population,count)) +
  geom_smooth(method=lm, forumla =count~Population,se=T,col="red",fill="orchid")+
  xlim(0,1000000)+
  ggpubr::theme_pubr()+
  xlab("Population")+
  ylab("Number of cases")+
  theme(axis.text.x = element_text(size=16))+
  theme(axis.title.x = element_text(size=16))+
  theme(axis.text.y = element_text(size=16))+
  theme(axis.title.y = element_text(size=16))+
  labs(title="Relation between population and number of pertussis cases ")
### relation between population and cases of all diseases
ggplot(dem.dis2, aes(Population,count)) +
  geom_smooth(method=lm, forumla =count~Population,se=T,col="orange",fill="yellow")+
  xlim(0,1000000)+
  ggpubr::theme_pubr()+
  xlab("Population")+
  ylab("Number of cases")+
  theme(axis.text.x = element_text(size=16))+
  theme(axis.title.x = element_text(size=16))+
  theme(axis.text.y = element_text(size=16))+
  theme(axis.title.y = element_text(size=16))+
  labs(title="Relation between population and number of all disease cases ")
```
```{r}
#B. Density
###################densiyt
### relation between population and cases of pertusis
ggplot(de.dis.2010.per, aes(popdens,count)) +
  geom_smooth(method=lm, forumla =count~popdens,se=T,col="purple",fill="light green")+
  #xlim(0,1000000)+
  ggpubr::theme_pubr()+
  xlab("Population density")+
  ylab("Number of cases")+
  theme(axis.text.x = element_text(size=16))+
  theme(axis.title.x = element_text(size=16))+
  theme(axis.text.y = element_text(size=16))+
  theme(axis.title.y = element_text(size=16))+
  labs(title="Relation between population density and number of pertussis cases ")
### relation between population and cases of all diseases
ggplot(dem.dis2, aes(popdens,count)) +
  geom_smooth(method=lm, forumla =count~podens,se=T,col="navy blue",fill="ligh tblue")+
  #xlim(0,1000000)+
  ggpubr::theme_pubr()+
  xlab("Population density")+
  ylab("Number of cases")+
  theme(axis.text.x = element_text(size=16))+
  theme(axis.title.x = element_text(size=16))+
  theme(axis.text.y = element_text(size=16))+
  theme(axis.title.y = element_text(size=16))+
  labs(title="Relation between population and number of all disease cases ")

```

